Predictive Quantization for MIMO-OFDM SVD Precoders using Reservoir Computing Framework

### 2019-04-19 paper/main.tex
Changes in Section5

The output of the reservoir is $\bp_{i,t}$, a vectorized representation of $\bP_{i,t}$.
Therefore, the output of the reservoir is also a $D_{\text{in}}$ dimensional vector ($D_{\text{out}}=D_{\text{in}}$).
The output $\bp_{i,t}$ is obtained via matrix transformation of the reservoir state $\br_{i,t}$ via the $D_{\text{resv}}\times D_{\text{out}}$ output coupler matrix $\bW^{\text{out}}_{i,t-1}$\footnote{Why it is $t-1$ and not $t$ would surface up in Section \ref{train}}, viz.
\begin{equation}
\bp_{i,t}=\bW^{\text{out}}_{i,t-1}\br_{i,t}
\label{pred_op}
\end{equation}
\begin{figure*}[ht]
\centering
\includegraphics[width=0.8\textwidth]{images/system.pdf}
\caption{An overview of the reservoir computing framework proposed for predictive quantization of $\bu_{i,t}$}
\label{res_overview}
\end{figure*}
This completes the forward prediction to obtain $\bP_{i,t}$ using only the past observed $\bU_{i,t-1}$ (i.e. $p=1$, Section \ref{section2}). That is,
\begin{equation}
\bP_{i,t}=f_P(\tilde{\bU}_{i,t-1})=\bW^{\text{out}}_{i,t-1}\text{tanh}(\bA_i\br_{i,t-1}+\bW^{\text{in}}_i\tilde{\bu}_{i,t-1})\footnote{For brevity, $\bp_{i,t}$ and $\bP_{i,t}$ are treated to be one and the same}
\label{fwd_pred}
\end{equation}
However, a complete discussion of the prediction algorithm also involves a backward pass, involving the updation of output couplers $\bW^{\text{out}}_{i,t}$, which would follow subsequently.
\subsection{Quantization function $q_P$}
Before proceeding, it is critical to describe the quantization function $q_P(\bP_{i,t},\text{fb}_{i,t})$, which is similar to the one used in \cite{6891198,Gupt1905:Predictive,6545375}.
Central to the quantization algorithm is the fact that tangent spaces local to a point in manifold are linear vector spaces.
Given two points $\bX$ and $\bY$ in $\text{St}(N_T,N_R)$, a lifting operation $\bT^Y_X=\texttt{lift}(\bX,\bY)$, $\texttt{lift}:\text{St}(N_T,N_R)\times \text{St}(N_T,N_R) \to \mathcal{T}_X \text{St}(N_T,N_R)$ yields a tangent from $X$ to $Y$, denoted as $\bT^Y_X \in \mathcal{T}_X \text{St}(N_T,N_R)$, where $\mathcal{T}_X \text{St}(N_T,N_R)$ is the local tangent space at $\bX$.
A corresponding retraction operation $Y=\texttt{retract}(X,\bT^Y_X)$, $\texttt{retract}: \text{St}(N_T,N_R)\times \mathcal{T}_X \text{St}(N_T,N_R) \to \text{St}(N_T,N_R)$ gives back the manifold point obtained by traversing in the tangent direction given by the second argument.
In this work, the chosen lifting-retraction operations are the Cayley Exponentials, discussed in \cite{DBLP:journals/corr/abs-1708-00045,Gupt1905:Predictive}. % and used in \cite{Gupt1905:Predictive}.
The Cayley exponential lifting operation maps two points in $\text{St}(N_T,N_R)$, to a $N_T\times N_T$ skew Hermitian matrix that represents the tangent from first point to the other.

@@

This completes our technical discussion for the complete reservoir computing based framework for predictive quantization illustrated in Fig. %Ref Fig.

@@

\section{Simulation Results}
\label{section4}
\subsection{Simulation setting considered}
\label{setting}
\noindent We simulate both our predictive quantization algorithms and the one presented in \cite{6891198} for the IEEE Pedestrian-A channel, with $N_T=4$, $N_R=2$. We simulated our algorithms by using the IT++ library through the Python wrapper py-itpp \cite{ViditPy}, with the channels being generated using Jake’s model for realistic simulations. Although in both our work and \cite{6891198}, only temporal correlations are exploited, we compare the BER, quantization error and logsum rate results for the MIMO-OFDM setting with number of subcarriers $N=64$, and evenly spaced $8$ feedback subcarriers, indexed from 0, viz. $\{0,9,\ldots,63\}$ ($\{9k, k \in \{0,1,\ldots,7\}\}$). The precoding matrices at non fedback subcarriers ($i\neq\{9k\}$) are estimated via the Cayley Interpolation method presented in \cite{Gupt1905:Predictive}. Feedback is available for $i=\{9k\}$ subcarriers (i.e. transmitter obtains $\text{fb}_{\{9k\},t}$) for predictive quantization. The setting is visually illustrated in Fig. \ref{table}. A 6 bit codebook generated via Lloyd codebook algorithm for the Stiefel Manifold in \cite{6678348} is used to feedback $\tilde{\bU}_{\{9k\},0}$ and differential quantization (i.e. $\bP_{i,t}=\tilde{\bU}_{i,t-1}$) is used for 10 time instances to provide provide the initial training data for the reservoir and for overcoming initial transient response of the reservoir discussed in \cite{mosleh2017brain}. A 6 bit base codebook is used for the quantization function $q_P$, generated via vector quantization via $k$-means over a tangent list, comprising of actual tangents from the predicted precoder to the actual precoder obtained from SVD of channel matrix, to guarantee isotropic codewords \cite{Gupt1905:Predictive}.

For prediction of each of the $8$ fed-back subcarriers, we consider $8$ separate reservoirs, each of them evolved separately, with $i\in\{9k, k \in \{0,1,\ldots,7\}\}$ in both the forward pass prediction and quantization ($f_P,q_P$) \eqref{fpqp} and backward pass to update $\bW_{i,t}^{\text{out}}$ \eqref{backpass}. Unlike \cite{Gupt1905:Predictive}, we don't exploit frequency correlations for predictive quantization, and the $8$ reservoirs for prediction of $8$ fed-back subcarriers are kept independent of each other. All these 8 resrvoirs are initialized with $\bA_{i\in\{9k\}}$ being an Erdos-Renyi\" graph, with probability of edge connections being $0.2$, $\lambda=2$, $\bW^{\text{out/in}}_{i\in\{9k\},0}$ being randomly assigned, $D_{\text{in}}=D_{\text{out}}=14$ (vectorized representation of $4\times2$ semi-unitary matrix, Section \ref{vectorize}) and $D_{\text{resv}}=60$.

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{images/qtizErr.pdf}
\caption{Quantization Error in terms of chordal distance to the actual precoder $\bU_{\{9k\},t}$ observed at the receiver, from the quantized estimate $\tilde{\bU}_{\{9k\},t}$ at the transmitter, averaged over 8 feedback subcarriers ($\{9k, k \in \{0,1,\ldots,7\}\}$) and 10 independent channel realizations ($f_DT_s=3.5\times10^{-4}$)}
\label{qtiz_err}
\end{figure}

>>>>

The output of the reservoir is $\bp_{i,t}$, a vectorized representation of $\bP_{i,t}$.
Therefore, the output of the reservoir is also a $D_{\text{in}}$ dimensional vector ($D_{\text{out}}=D_{\text{in}}$).
The output $\bp_{i,t}$ is obtained via matrix transformation of the reservoir state $\br_{i,t}$ via the $D_{\text{resv}}\times D_{\text{out}}$ output coupler matrix $\bW^{\text{out}}_{i,t-1}$\footnote{Why it is $t-1$ and not $t$ would surface up in Section \ref{train}}, viz.
\begin{equation}
\bp_{i,t}=\bW^{\text{out}}_{i,t-1}\br_{i,t}
\label{pred_op}
\end{equation}
\begin{figure*}[ht]
\centering
\includegraphics[width=0.8\textwidth]{images/system.pdf}
\caption{An overview of the reservoir computing framework proposed for predictive quantization of $\bu_{i,t}$}
\label{res_overview}
\end{figure*}
This completes the forward prediction to obtain $\bP_{i,t}$ using only the past observed $\bU_{i,t-1}$ (i.e. $p=1$, Section \ref{section2}). That is,
\begin{equation}
\bP_{i,t}=f_P(\tilde{\bU}_{i,t-1})=\bW^{\text{out}}_{i,t-1}\text{tanh}(\bA_i\br_{i,t-1}+\bW^{\text{in}}_i\tilde{\bu}_{i,t-1})\footnote{For brevity, $\bp_{i,t}$ and $\bP_{i,t}$ are treated to be one and the same}
\label{fwd_pred}
\end{equation}
However, a complete discussion of the prediction algorithm also involves a backward pass, involving the updation of output couplers $\bW^{\text{out}}_{i,t}$, which would follow subsequently.
\subsection{Quantization function $q_P$}
\label{qp}
Before proceeding, it is critical to describe the quantization function $q_P(\bP_{i,t},\text{fb}_{i,t})$, which is similar to the one used in \cite{6891198,Gupt1905:Predictive,6545375}.
Central to the quantization algorithm is the fact that tangent spaces local to a point in manifold are linear vector spaces.
Given two points $\bX$ and $\bY$ in $\text{St}(N_T,N_R)$, a lifting operation $\bT^Y_X=\texttt{lift}(\bX,\bY)$, $\texttt{lift}:\text{St}(N_T,N_R)\times \text{St}(N_T,N_R) \to \mathcal{T}_X \text{St}(N_T,N_R)$ yields a tangent from $X$ to $Y$, denoted as $\bT^Y_X \in \mathcal{T}_X \text{St}(N_T,N_R)$, where $\mathcal{T}_X \text{St}(N_T,N_R)$ is the local tangent space at $\bX$.
A corresponding retraction operation $Y=\texttt{retract}(X,\bT^Y_X)$, $\texttt{retract}: \text{St}(N_T,N_R)\times \mathcal{T}_X \text{St}(N_T,N_R) \to \text{St}(N_T,N_R)$ gives back the manifold point obtained by traversing in the tangent direction given by the second argument.
In this work, the chosen lifting-retraction operations are the Cayley Exponentials, discussed in \cite{DBLP:journals/corr/abs-1708-00045,Gupt1905:Predictive}. % and used in \cite{Gupt1905:Predictive}.
The Cayley exponential lifting operation maps two points in $\text{St}(N_T,N_R)$, to a $N_T\times N_T$ skew Hermitian matrix that represents the tangent from first point to the other.

@@

This completes our technical discussion for the complete reservoir computing based framework for predictive quantization illustrated in Fig. \ref{res_overview}

@@

\section{Simulation Results}
\label{section4}
\subsection{Simulation setting considered}
\label{setting}
\noindent We simulate both our predictive quantization algorithms and the one presented in \cite{6891198} for the IEEE Pedestrian-A channel, with $N_T=4$, $N_R=2$. 
The channel matrices $\bH_{i,t}$ are generated using Jake’s model for realistic simulations, by using the IT++ library through the Python wrapper py-itpp \cite{ViditPy}. 
Although in both our work and \cite{6891198}, only temporal correlations are exploited, we compare the BER, quantization error and logsum rate results for the complete MIMO-OFDM setting with number of subcarriers $N=64$, and evenly spaced $8$ feedback subcarriers, indexed from 0, viz. $\{0,9,\ldots,63\}$ ($\{9k, k \in \{0,1,\ldots,7\}\}$). The precoding matrices at non fedback subcarriers ($i\neq\{9k\}$) are estimated via the Cayley Interpolation method presented in \cite{Gupt1905:Predictive}. Feedback is available for $i=\{9k\}$ subcarriers (i.e. transmitter obtains $\text{fb}_{\{9k\},t}$) for predictive quantization. The setting is visually illustrated in Fig. \ref{table}. A 6 bit codebook generated via Lloyd codebook algorithm for the Stiefel Manifold in \cite{6678348} is used to feedback $\tilde{\bU}_{\{9k\},0}$ and differential quantization (i.e. $\bP_{i,t}=\tilde{\bU}_{i,t-1}$) is used for 10 time instances to provide provide the initial training data for the reservoir and for overcoming initial transient response of the reservoir discussed in \cite{mosleh2017brain}. A 6 bit base codebook is used for the quantization function $q_P$, generated according to the method presented in Section \ref{qp} and \cite{Gupt1905:Predictive}. Hence, for both initialization and subsequent feedback, $\text{fb}_{i,t}$ can be encoded via 6 bits (i.e. $\mathcal{B}=6$ Section \ref{section2}). The bit budget per OFDM frame is thus 48 bits, for total 8 equally spaced feedback subcarriers $\{9k, k \in \{0,1,\ldots,7\}\}$.

For prediction of each of the $8$ fed-back subcarriers, we consider $8$ separate reservoirs, each of them evolved separately, with $i\in\{9k, k \in \{0,1,\ldots,7\}\}$ in both the forward pass prediction and quantization ($f_P,q_P$) \eqref{fpqp} and backward pass to update $\bW_{i,t}^{\text{out}}$ \eqref{backpass}. Unlike \cite{Gupt1905:Predictive}, we don't exploit frequency correlations for predictive quantization, and the $8$ reservoirs for prediction of $8$ fed-back subcarriers are kept independent of each other. All these 8 resrvoirs are initialized with $\bA_{i\in\{9k\}}$ being an Erdos-Reny\"i graph, with probability of edge connections being $0.2$, $\lambda=2$, $\bW^{\text{out/in}}_{i\in\{9k\},0}$ being randomly assigned, $D_{\text{in}}=D_{\text{out}}=14$ (vectorized representation of $4\times2$ semi-unitary matrix, Section \ref{vectorize}) and $D_{\text{resv}}=60$.

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.45\textwidth]{images/qtizErr.pdf}
\caption{Quantization Error in terms of chordal distance to the actual precoder $\bU_{\{9k\},t}$ observed at the receiver, from the quantized estimate $\tilde{\bU}_{\{9k\},t}$ at the transmitter, averaged over 8 feedback subcarriers ($\{9k, k \in \{0,1,\ldots,7\}\}$) and 10 independent channel realizations ($f_DT_s=3.5\times10^{-4}$)}
\label{qtiz_err}
\end{figure}
